---
title: "EDA"
author: "Francois van Zyl"
date: "13/09/2021"
output: 
  html_document: 
    theme: lumen
---


```{r warning=F, message=F}
# Libraries ---------------------------------------------------------------
library(ggplot2)
library(knitr)
library(dplyr)
library(kableExtra)
library(autoEDA)
library(rattle)
library(rpart)
library(rpart.plot)
library(partykit)
library(tidyrules)
library(papeR) # For formatting
# Separate numerical + categorical ----------------------------------------
setwd('C://Users//jd-vz//Desktop//Code//src//explore//')
df <- read.csv(file = 'C://Users//jd-vz//Desktop//Code//data//collected_fpl.csv')
# df <- read.csv(file = 'C://Users//jd-vz//Desktop//Code//data//collected_us.csv')
num_feat <- df[sapply(df, is.numeric)]
cat_feat <- df[,colnames(df)[grepl('factor|logical|character',sapply(df,class))]]


```



```{r }
# These features could be interesting to investigate
unique(df$clean_sheets)
unique(df$clean_sheets)
unique(df$penalties_missed) 
unique(df$penalties_saved) # This is not a binary feature
unique(df$yellow_cards)
unique(df$red_cards)
unique(df$was_home)
```
# By position

```{r}
df_temp <- df
pos_test <- summarize(df_temp, type = "numeric", group = 'position', test = FALSE, quantiles = TRUE)
pos_test %>% kable() %>% kable_styling(full_width = FALSE)
```
## Observations 
* Assists
  + GK have a lower spread of assists
  + A GK scored an assist, potential outlier
  
* Bonus
  + FWD have a significantly higher mean and SD than the rest

* BPS
  + FWD have the highest SD 
  + FWD have the lowest mean
  + DEF have the highest mean
  
* Creativity
  + GK has a very low mean and SD
  + MID have the highest creativity.
  
* Goals Scored
  + GK scored a goal, potential outlier
  + FWD score more goals
  + MID scored four goals
  
* ICT
  + FWD has highest mean and SD
  + MID has max ICT


* Influence
  + Def the highest avg influence, but not close to others max.

* minutes
  + On avg, GK plays the least and DEF the most

* Own goal
  + DEF do it the most, FWD the least

* Penalties missed
  + DEF. FWD. and MID have taken shots and missed penalties.

* Red cards
  + Defenders get the most, only marginally
  + Goalkeepers get the least

* Saves
  + GK specific

* Selected
  + As expected, more FWD are selected
  
* Threat
  + GK low.
  + Mid surprisingly high
  
* Total points
  + GK score least
  + FWD score most
  
* Value
  + GK and DEF have a lower average value
  + FWD and MID more on avg.

* Yellow cards
  + More DEF and MID get yellow cards

# By home (with p-value)

```{r}
df_temp <- df
pos_test <- summarize(df_temp, type = "numeric", group = 'was_home', 
                      test = c("wilcox.test"), quantiles = TRUE) # Since normality cannot be assumed
pos_test %>% kable() %>% kable_styling(full_width = FALSE) 
```

## Observations 

* BPS
  + More BPS obtained at home (slight)

* Clean sheets
  + More clean sheets are obtained at home

* Creativity
  + More creativity at home


* Goals conceded
  + Less goals are conceded at home

* ICT
  + A higher ICT index is obtained at home

* Penalties missed
  + More penalties missed away from home

* Penalties saved
  + More penalties saved at home

* Threat
  + A higher threat is observed at home

* Total points
  + More total points scored at home

* Transfer balance
  + Negative balance at home, positive away from home

# Clean sheets

```{r }
df_temp <- df
df_temp$clean_sheets <- as.factor(df_temp$clean_sheets)
cs_test <- summarize(df_temp, type = "numeric", group = 'clean_sheets', test = c("wilcox.test"))
cs_test %>% kable() %>% kable_styling(full_width = FALSE,) 
```

## Observations

Sample size? 42028 vs 4839a


# Penalties missed

```{r }
df_temp <- df
df_temp$penalties_missed <- as.factor(df_temp$penalties_missed)
pm_test <- summarize(df_temp, type = "numeric", group = 'penalties_missed', test = c("wilcox.test"))
pm_test %>% kable() %>% kable_styling(full_width = FALSE) 
```

## Observations

Sample size? 46824 vs 43 



# Red cards

```{r }
df_temp <- df
df_temp$red_cards <- as.factor(df_temp$red_cards)
rc_test <- summarize(df_temp, type = "numeric", group = 'red_cards', test = c("wilcox.test"))
rc_test %>% kable() %>% kable_styling(full_width = FALSE) 
```

## Observations

Sample size? 46774 vs 93


# Yellow cards

```{r }
df_temp <- df
df_temp$yellow_cards <- as.factor(df_temp$yellow_cards)
yc_test <- summarize(df_temp, type = "numeric", group = 'yellow_cards', test = c("wilcox.test"))
yc_test %>% kable() %>% kable_styling(full_width = FALSE) 
```

## Observations

Sample size? 44494 vs 2373





# Investigate

```{r}
summarize(df, type = "numeric", group = NULL,
                       test = FALSE, count = FALSE)  %>% kable() %>% kable_styling(full_width = FALSE) 
```

## Observations 

* BPS, creativity, influence, threat, ict index, very skew.

* YC, value transfers, total_points, skew

* 


* 


* 


* 


* 


* 


* Strengths seem normal





* 
  + 


* 
  + 


* 
  + 


* 
  + 


* 
  + 


* 
  + 


* 
  + 






